---
title: "K-12 to MELCs Curriculum"
subtitle: "Initial Insights"
author: "Shaine Rosewel P. Matala"
output:
  html_document:
    df_print: paged
    theme: united
    toc: true
    toc_folding: yes
  html_notebook:
    df_print: paged
    theme: united
---

```{r echo = FALSE, message=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.align = "center", fig.width=6, fig.height=6)
library(readxl)
library(dplyr)
library(magrittr)
```

```{r}
# get ll relevant sheets
sheet.all <- 
lapply(2:12, function(.){
  read_xlsx("INIT_01.xlsx", sheet = ., trim_ws = TRUE, range = cell_cols("A:J"))
})

# gather all sheets into one dataframe
init_01 <- do.call("rbind", sheet.all)

# remove extra white spaces within
init_01 <- as.data.frame(apply(init_01, 2, stringr::str_squish), row.names = FALSE)
```

## At a Glance

```{r tidy the data}
library(dplyr)
tidy.1 <- init_01[!is.na(init_01$MELC_LC),]
tidy.2 <- tidyr::fill(tidy.1, c(QUARTER, WEEK))
CURRICULUM <- rep("MELC", nrow(tidy.2))
curr.1 <- cbind( CURRICULUM, tidy.2[,c(1,2,5:6)])
colnames(curr.1)[4:5] <- c("LC", "CODE")

CURRICULUM <- rep("K12", nrow(init_01))
curr.2 <- cbind(CURRICULUM, init_01[,c(1,2,3,4)])
colnames(curr.2)[4:5] <- c("LC", "CODE")

curr <- rbind(curr.2, curr.1)
curr[,1:3] <- lapply(curr[,1:3], as.factor)
```

At first glance, it is easy to say that the K-12 curriculum has been trimmed down. The question of whether the reduction is meaningful and significant is yet to be answered. By observation, the difference in the number of learning competencies from Kinder to Grade 6 are generally higher than from those that remain (high school, in particular). Note that for plotting, DepEd provided codes were used as bases.

```{r prelim plot1, fig.width=8, fig.align = "center"}
curr.3 <- curr %>% dplyr::group_by(CURRICULUM, GRADE) %>% tally()
library(ggplot2)
ggplot(curr.3, aes(x = GRADE, y = n, fill = CURRICULUM)) +
  geom_col(position = "dodge") +
  geom_text(aes(label=n), position=position_dodge(width=0.9), vjust=-0.5) +
  ggtitle("Number of Math Competencies Per Grade") +
  scale_x_discrete(labels=c("K", "1", "2", "3", "4", "5", "6", "7", "8", "9", "10")) +
  xlab("Grade") +
  ylab(NULL) +
  theme(axis.text.y=element_blank(),
        axis.ticks.y=element_blank()) 
```

The curriculum is supposed to be spiral. In our context, this is how spiraling looks like, according to DepEd. The question is, is it supposed to be really like this? Guided by the principles and theories in education, do these panels suggest that we have a nicely  weaved curriculum? Where is Stats and Prob in grades 8 and 9 when it is present in grades 7 and 10? Are we not mistaken? Is there really a gap? (In fact there is an error in the codes presented by DepEd. Read further ([Wrong codes] section) to see. Grade 8 has stats and prob. But still, grade 9 manifests the gap.)

```{r prelim plot2, fig.width=8, fig.align = "center"}
curr.3 <- curr %>% dplyr::group_by(CURRICULUM, GRADE, `CONTENT/STRAND`) %>% tally()# %>% filter(GRADE %in% c("00", "01", "02", "03", "04", "05", "06"))
curr.3$`CONTENT/STRAND` <- factor(curr.3$`CONTENT/STRAND`, levels=c("LG", "SP", "ME", "AL", "GE", "NS"))
ggplot(curr.3, aes(x = GRADE, y =as.numeric(n), fill=`CONTENT/STRAND`)) +
  geom_col(position = "fill") +
  scale_fill_brewer(type = "qual", palette = "Set1")+
  ggtitle("How Content Progresses From K-10") +
  xlab(NULL) +
  ylab(NULL) +
  theme(axis.text.y=element_blank(),
        axis.ticks.y=element_blank()) +
  facet_grid(rows=vars(CURRICULUM)) +
  scale_x_discrete(labels=c("K", "1", "2", "3", "4", "5", "6", "7", "8", "9", "10")) 
 #theme(legend.position="bottom")
```
```{r}
library(DT)
library(rio)
datatable(init_01[, 1:6], rownames = FALSE, filter = 'top')
```
The question we are dying to address: Was there really a reduction in the number of learning competencies? Let us have a real quick view.


## Problems to begin with

### Duplicated LC

There is one `K12_LC` in the K12 curriculum that is duplicated. Notice that they are differently coded, too. (hit the right arrow to see)

```{r}
dat <- dplyr::filter(init_01, K12_LC %in% init_01$K12_LC[which(duplicated(init_01$K12_LC))])[11:12,3:4]
rownames(dat) <- NULL
dat
```

This was actually addressed in MELC by writing a `MELC_LC`, which I coded as `NEW` just because I am not sure to which it is exactly is equivalent to in the K12 curriculum.

```{r}
dat <- dplyr::filter(init_01, K12_LC %in% init_01$K12_LC[which(duplicated(init_01$K12_LC))])[11,5:6]
rownames(dat) <- NULL
dat
```

### Duplicated CODES

<span style="color: red;"> UPDATE (12-15-2020): More duplicated codes found in K12 Kinder. </span>

```{r}
dat <- dplyr::filter(init_01, K12_CODE %in% init_01$K12_CODE[which(duplicated(init_01$K12_CODE))])
rownames(dat) <- NULL
dat
```



In the manual checking of MELC, one `MELC_LC` was observed to be duplicated as it appears in two different quarters. Not sure if DepEd really wants it this way.

```{r}
dat <- init_01[which(nchar(init_01$QUARTER) > 2), 5:8]
rownames(dat) <- NULL
dat
```

### Wrong codes

One strand in G8 is wrongly coded. It would appear that the codes are not 100% reliable. This error has been radiated to MELC as well. My wish is that for DepEd to be transparent on the information on how the codes were assigned by providing a short discussion about it within the curriculum, so that the meanings will make sense and allow errors such as this to be reported ASAP by curriculum users.

```{r}
prob <- (init_01 %>% filter(GRADE == "08", `CONTENT/STRAND`=="GE"))[20:24,3:6]
rownames(prob) <- NULL
prob
```


## Changes as they say

What I did is to compare the competencies from the K12 curriculum to those in the MELC curriculum and to indicate in the `MELC_CODE` which `K12_CODES` were merged to form the new `MELC_LC` `_` (underscore) was used to separate these merged `K12_CODE`. The worst case is that 6 `K12_CODE`s were all combined into one `MELC_LC`. This is not simply merging, in defense to DepEd. What they did is 

1. reduce verbs  
    ex: "visualizes, represents, and compares" changed to "compares"
2. separated but related competencies were combined in one grand competency  
    ex: M1NS-IIa-26.1_M1NS-IIa-28.1a_M1NS-IIb-26.2_M1NS-IIb-27.1_M1NS-IIc-27.2_M1NS-IIc-27.3
    
> add some more

```{r}
new.or.merged.melc <- init_01$MELC_CODE[grep(pattern = "(.+_.+)|(NEW)", x = init_01$MELC_CODE)]
data.frame(new.or.merged.melc)
```

Feel free to explore below the reduction that we are being told about. This shows the merging.

```{r}
merged.melc <- init_01$MELC_CODE[grep(pattern = ".+_.+", x = init_01$MELC_CODE)]
MELC2 <- unlist(strsplit(merged.melc, split = "_"))
merged.melc.huh <- dplyr::filter(init_01, K12_CODE %in% MELC2)[,c(1, 3:6)]
datatable(merged.melc.huh, rownames = FALSE, filter = 'top')
```

This shows the newly stated `MELC_LC`. I coded these as `NEW` because DepEd did not code them with a `K12_CODE` and so I assume that these are either 1. new competencies or 2. a mix of many `K12_LC`s. However, I can't exactly point out which ones were merged. It appears that 2nd there could be coded similarly to how it was in `K12_CODE`. I am not sure tho because DepEd only used `K12_CODE` for exact matches only.

```{r}
new.melc <- which(init_01$MELC_CODE == "NEW")
datatable(init_01[new.melc,3:6], options = list(paging = FALSE), rownames = NULL)
```

