---
title: "Initial Insights"
author: Shaine Rosewel P. Matala
output: 
  html_notebook:
    df_print: paged
---

```{r echo = FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## The Curriculums

Here are the column names for your reference: `r colnames(init_01)`

```{r}
# get ll relevant sheets
sheet.all <- 
lapply(2:12, function(.){
  read_xlsx("INIT_01.xlsx", sheet = ., trim_ws = TRUE, range = cell_cols("A:J"))
})

# gather all sheets into one dataframe
init_01 <- do.call("rbind", sheet.all)

# remove extra white spaces within
init_01 <- as.data.frame(apply(init_01, 2, stringr::str_squish))
init_01
```



## Duplicates

There is one `K12_LC` in the K12 curriculum that is duplicated. Notice that the codes are different as well. Click the right arrow to see that this was somehow addressed in the MELC by writing a new `MELC_LC`.

```{r}
dplyr::filter(init_01, K12_LC %in% init_01$K12_LC[which(duplicated(init_01$K12_LC))])[11:12,3:6]
```

In the manual checking of MELC, one `MELC_LC` was observed to be duplicated as it appears in two different quarters

```{r}
init_01[which(nchar(init_01$QUARTER) > 2), 5:8]
```



