---
title: "R Notebook"
output: html_notebook
---

# Project structure

An `R project` with `raw_curr` folder where the downloaded curriculum will be stored and `forprocessing` folder where the subset curriculum and .xlsx conversions will be stored

# Download curriculum files

Download the K12 curriculum

```{r}
#dir.create("raw_curr")
path <- "C:/Users/shain/Desktop/CURRICULUM_STUDIES/raw_curr/"
url2<-"https://www.deped.gov.ph/wp-content/uploads/2019/01/Math-CG_with-tagged-math-equipment.pdf"
destfile2 <- paste0(path, "K12_math_curr.pdf")
download.file(url2, destfile2, mode="wb")

urlk1 <- "https://www.deped.gov.ph/wp-content/uploads/2019/01/Kinder-CG_0.pdf"
destfile3 <- paste0(path, "K12K_math_curr.pdf")
download.file(urlk1, destfile3, mode="wb")
```

Download the MELC curriculum. Change the path as needed.

```{r}
url1<-"https://commons.deped.gov.ph/melc_k12"
destfile1 <- paste0(path, "MELC_math_curr.pdf")
download.file(url1, destfile1, mode="wb")
```



# Extract relevant pages and save

Extract relevant pages: page 9 - 253 of the K12 curriculum.

```{r}
#dir.create("forprocessing")
pdftools::pdf_subset("raw_curr/K12_math_curr.pdf", pages=9:253, output="forprocessing/K12_math.pdf")
pdftools::pdf_subset("raw_curr/K12K_math_curr.pdf", pages=17:21, output="forprocessing/K12K_math.pdf")
```

Extract relevant pages: page 197 - 240 of the MELC curriculum (only the math part)

```{r}
pdftools::pdf_subset("raw_curr/MELC_math_curr.pdf", pages=197:240, output="forprocessing/melc_math.pdf")
```

# Convert to .xlsx file

Go to [PD2GO]("https://www.pdf2go.com/pdf-to-excel")

# Tidying 

This is a scratch

## Curriculum files

Read the data and separately process elementary from jhs curriculum since reading them altogether makes us miss a lot. Maybe the problem is with the formatting

```{r}
k12<-readxl::read_xlsx("forprocessing/K12_math.xlsx")
k12_elem<-readxl::read_xlsx("forprocessing/K12_math.xlsx", skip=3, n_max=1102, col_names = TRUE)
k12_jhs<-readxl::read_xlsx("forprocessing/K12_math.xlsx", skip=1107, col_names = TRUE)
k12K<-readxl::read_xlsx("forprocessing/K12K_math.xlsx", skip=7, n_max=70)
```


Plan is to later on merge `k12_elem` and `k12_jhs` again. Hence, immediate objective is to figure out weirdly named columns and make them uniform on the mentioned datasets

Remove all columns that are all `NA`

```{r}
k12 <- list(k12_elem, k12_jhs, k12K)

k12_wo_na_cols <-
lapply(k12, function(x){
  x[, colSums(is.na(x)) != nrow(x)]
})
#Checks that codes do the right thing
# View(k12_jhs[,which(colSums(is.na(k12_jhs)) == nrow(k12_jhs))])
# k12_elem[, colSums(is.na(k12_elem)) != nrow(k12_elem)]
# 
# lapply(k12_wo_na_cols, dim)


```

Figure out colums with weird names

```{r}
k12_elem <- k12_wo_na_cols[[1]]
k12_jhs <- k12_wo_na_cols[[2]]
k12_k <- k12_wo_na_cols[[3]]
```


```{r}
library(magrittr)
library(dplyr)

weird_name_checker <- function(curr=k12_elem){
  weird_index <- grep(pattern="...[0-9]+", x=colnames(curr))
  weird_colname <- colnames(k12_elem)[weird_index]
  apply(curr[, weird_index], 2, function(x){
  #which.max returns index ng may highest frequency, names gives back the name!
  names(which.max(table(x)))
  })
  }
  
View(weird_name_checker(k12_k))
#weird_name_checker(k12_jhs)
```




Next is `coalesce` what should go together.

```{r}
library(doParallel)
settle_weird_colnames <-
function(curr_prepro="elem", this="whole") {
  
  if (curr_prepro == "elem") {
    curr_wo_na = k12_elem
  } else if (curr_prepro == "jhs") {
    curr_wo_na = k12_jhs
  } else if (curr_prepro == "k") {
    curr_wo_na = k12_k
  } else {
    print("mali input mo")
  }
  
  x<-names(table(weird_name_checker(curr_wo_na)))
  
  oo<-foreach(i=x, .combine=cbind) %do% {
    k <-append(curr_wo_na[, which(colnames(curr_wo_na)==i)],
        as.list(curr_wo_na[,names(which(weird_name_checker(curr_wo_na)==i))]))
    k1=coalesce(!!!k) #coalesce a list of vectors
    }
  
  colnames(oo) <- x
  a<- setdiff(colnames(curr_wo_na)[-(grep(pattern = "...[0-9]+" , x = colnames(curr_wo_na)))],
              colnames(oo))
  b<-which(colnames(curr_wo_na)%in% a)
    fin<-cbind(curr_wo_na[,b], oo)
  
  if (curr_prepro %in% c("elem", "jhs")) {
    col_order <- union(colnames(curr_wo_na)[-(grep(pattern = "...[0-9]+" , x = colnames(curr_wo_na)))],colnames(oo))
    fin2 <- fin[, col_order]
  } else if (curr_prepro == "k") {
    colnames(fin)[c(1,2,4,6)] <- c("CONTENT STANDARDS", "PERFORMANCE STANDARDS", "LEARNING COMPETENCY", "CONTENT")
    fin2 <- fin[, c(6, 1, 2, 4, 3, 5)]
  }
  
  fin3 <- anti_join(fin2, fin2[which(duplicated(fin2)),])
  
  #=============================================================================
  
  
o=gsub("(NA)|(\r\n)|( +)", x=paste(fin3$CODE, collapse = "_"), replacement = "")
l=paste(fin3$`LEARNING COMPETENCY`, collapse = "_")
    
if(curr_prepro=="elem"){
  c_1 = stringr::str_extract_all(o, "[A-Z][0-9]+[A-Z]+-[I,V]+[a-z]-[0-9]+(\\.[0-9]+)?")[[1]]
  } else if (curr_prepro=="jhs") {
  c_1 = stringr::str_extract_all(o, "[A-Z][0-9]+[A-Z]+-[I,V]+[a-z]-([0-9]|[a-z])")[[1]]
  } else if (curr_prepro=="k") {
  c_1 = stringr::str_extract_all(o, "MK[A-Z]+(-|_00-[0-9]+")[[1]]
  } else {
  print('mali supplied mo')
  }
    
o1<-gsub("(NA)|(\r\n)", x=l, replacement = "")
o2<-gsub("(e\\.g\\.)|([1-9]+[0-9]*\\.[0-9]+)", x=o1, replacement = "-") 
o3<-gsub("10\\. +10\\.", x=o2, replacement = "10\\.")
o4<-regmatches(o3,gregexpr(text=o3, pattern = "[1-9]+[0-9]*\\.( |[a-zA-Z]).+?(_|\\.) *[a-z]*"))
lc = sapply(o4, function(.){gsub(pattern = "_+$", x = ., replacement = "")})
  
  # if(curr=="elem"){
  #   a=cbind(c_elem, lc)
  # } else if (curr=="jhs") {
  #   a=cbind(c_jhs, lc)
  # } else {
  #   print('select between "elem" and "jhs" for curr')
  # }
a1 <- cbind(c_1, lc)
colnames(a1) <- c("CODE", "LEARNING COMPETENCY")

if (this=="whole"){
  return(fin3)
} else if (this=="lc_code") {
  return(a1)
} else {
  print("walang ganyang supply choice")
}
}
```



```{r}
library(tibble)
#curr=settle_weird_colnames(k12_elem)
code_lc <- function(curr="elem"){
  if(curr=="elem"){
    o=gsub("(NA)|(\r\n)|( +)", x=paste(settle_weird_colnames(k12_elem)$CODE, collapse = "_"), replacement = "")
    l=paste(settle_weird_colnames(k12_elem)$`LEARNING COMPETENCY`, collapse = "_")
    c_elem = stringr::str_extract_all(o, "[A-Z][0-9]+[A-Z]+-[I,V]+[a-z]-[0-9]+(\\.[0-9]+)?")[[1]]
  } else if (curr=="jhs") {
    o=gsub("(NA)|(\r\n)|( +)", x=paste(settle_weird_colnames(k12_jhs)$CODE, collapse = "_"), replacement = "")
    l=paste(settle_weird_colnames(k12_jhs)$`LEARNING COMPETENCY`, collapse = "_")
    c_jhs = stringr::str_extract_all(o, "[A-Z][0-9]+[A-Z]+-[I,V]+[a-z]-([0-9]|[a-z])")[[1]]
  } else {
    print('select between "elem" and "jhs" for curr')
  }
  o1<-gsub("(NA)|(\r\n)", x=l, replacement = "")
  o2<-gsub("(e\\.g\\.)|([1-9]+[0-9]*\\.[0-9]+)", x=o1, replacement = "-") 
  o3<-gsub("10\\. +10\\.", x=o2, replacement = "10\\.")
  o4<-regmatches(o3,gregexpr(text=o3, pattern = "[1-9]+[0-9]*\\.( |[a-zA-Z]).+?(_|\\.) *[a-z]*"))
  lc = sapply(o4, function(.){gsub(pattern = "_+$", x = ., replacement = "")})
  
  if(curr=="elem"){
    a=cbind(c_elem, lc)
  } else if (curr=="jhs") {
    a=cbind(c_jhs, lc)
  } else {
    print('select between "elem" and "jhs" for curr')
  }
  colnames(a) <- c("CODE", "LEARNING COMPETENCY")
  return(a)
}

elem1<- as.tibble(code_lc(curr="elem"))
hs1<- as.tibble(code_lc(curr="jhs"))

elem2 <- elem1 %>% transform(Subject=substr(CODE, 1, 1),
                             Grade=substr(CODE, 2, 2),
                             Content=substr(CODE, 3, 4),
                             Quarter=str_extract(CODE, pattern="[I,V]+"),
                             what1 = str_extract(CODE, pattern="[a-z]+"),
                             what2 = str_extract(CODE, pattern="[0-9]+(\\.[0-9])?$")
                             )

k1=paste(elem2$Subject, elem2$Grade, elem2$Content, sep="") 
k2=paste(elem2$Quarter, sep="") #, elem2$what1, sep="") 
PSID=paste(k1, k2, sep="-")

elem3 <- cbind(elem2, PSID)
```


Use codes as ID

```{r}

elem <- at("elem", "whole")
library(tidyr)
a<-elem %>% filter(!is.na(`PERFORMANCE STANDARDS`) | !is.na(`CONTENT STANDARDS`)) %>%
  select(c(CODE, `CONTENT STANDARDS`, `PERFORMANCE STANDARDS`))%>% fill(CODE)


a1<-paste(a$CODE, collapse="")
a1<-gsub(a1, pattern=" ", replacement="")
a$CODE<-stringr::str_extract_all(a1, pattern="[A-Z][0-9]+[A-Z]+-[I,V]+[a-z]-[0-9]+(\\.[0-9]+)?")[[1]]

a <- a %>% mutate(PSID=stringr::str_extract_all(a1, pattern="[A-Z][0-9]+[A-Z]+-[I,V]+")[[1]])
View(a)
View(left_join(elem1, a, "CODE"))
```


PERFORMANCE STANDARDS

```{r}
library(stringr)
a2<-
a %>% group_by(PSID) %>% 
  summarise(PS =
              gsub("(NA)|(\r\n)", x=paste(`PERFORMANCE STANDARDS`, collapse = " "),
                   replacement = " ") %>%
              strsplit(split = "([1-9]+[0-9]*\\.)? *is able to") %>%
              unlist()) %>% 
  ungroup() %>%
  filter(PS!="")

a3<-
a %>% group_by(PSID) %>% 
  summarise(CS =
              gsub("(NA)|(\r\n)", x=paste(`CONTENT STANDARDS`, collapse = " "),
                   replacement = " ") %>%
              strsplit(split = "([1-9]+[0-9]*\\.)? *demonstrates") %>%
              unlist()) %>% 
  ungroup() %>%
  filter(CS!="")


a3_1<-a3 %>% group_by(PSID) %>% summarise(CS=paste(CS, collapse="_"))
a2_1<-a2 %>% group_by(PSID) %>% summarise(PS=paste(PS, collapse="_"))


elem4 <- left_join(elem3, a2_1, "PSID")
elem4$PS[duplicated(elem4$PS)]<-NA
elem4_1 <- left_join(elem4, a3_1, "PSID")
elem4_1$CS[duplicated(elem4_1$CS)]<-NA




```













